FraCaS: 80 percent complete

### 2019-02-27 paper2018/paper.tex
abstract added and shortened - see SC comments

\begin{abstract}
  TODO
\end{abstract}
\section{Introduction}
Natural Language Inference (NLI) is the task of determining of whether
an NL hypothesis H follows from an NL premise(s) P. NLI has received a
lot of attention in the Computational Semantics literature and has
been approached using a variety of techniques, ranging from logical
approaches
\citep{bos:2008,Mineshima:2015,Abzianidze:2015,bernardy_type_2017},
all the way to the recent Deep Learning (DL) models for NLI. The
latter approaches, following a general trend in NLP, have been
dominating NLI and a number of impressive results have been
produced \citep{kim:2018,radford:2018,liu:2019}.\footnote{These are
  the three systems with the best results on SNLI in increasing
  order at the time of writing.}  State-of-the-art DL systems achieve an accuracy of around
0.9 when tested on suitable datasets. However, the datasets that are
used are assuming a definition of inference that can be thought to be
`looser' or less precise compared to the definition assumed in
platforms based in logical approaches. Consider the following
two examples, one from the SNLI dataset, predominatly used to test DL
approaches and which has a state-of-the-art performace of 0.911
\citet{liu:2019}, and the FraCaS test suite, the dataset \textit{par
  excellence} to test logical approaches.
%
\begin{lingex}
\item % An SNLI example \\
  \label{ex:snli}
\textbf{P}  A  man selling donuts to a customer during a world
exhibition event held in the city of Angeles. \\
 \textbf{H}  A woman drinks her coffee in a small cafe.\\
\textbf{Label: Contradiction} [SNLI]
% \item % An example from FraCaS\\
% \textbf{P1} A Scandinavian won the Nobel Prize.\\
% \textbf{P2}	Every Swede is  Scandinavian.\\
% \textbf{Q}  Did a Swede win the Nobel prize? \\
% \textbf{H} A Swede won the Nobel prize.\\
% \textbf{Label: Yes} [FraCaS 173]
\end{lingex}
%
In \ref{ex:snli}, one has to make a number of non-trivial
assumptions in order to arrive at a contradiction: a) the two
situations described have to be taken to refer to the same situation
and, thus, decide whether one description contradicts the other, b)
the indefinite article in the premise has to be identified with the
indefinite article in the hypothesis. (Additionally considering that a
person cannot be a man selling donuts and a woman drinking coffee at
the same time.) While this can be part of the reasoning humans
perform, it is not the only possibility. More precise, logical
reasoning can be also performed. Such an example is the FraCaS
example, as well as very fine-grained cases like the Antecedent
Contained Deletion ellipsis example \ref{ex:acd} below.\todo{It'd be nice to have a more clear parallel with indefinite articles as well; but I could not find this in fracas.}

@@

\paragraph{GF}
In GF, abstract syntax is comprised of: a) a number of syntactic
categories, and b) a number of syntactic construction functions. The latter
provide the means to compose basic syntactic categories into more complex
ones.  For example, consider the constructor: $AdjCN : AP → CN → CN$. This  expresses
that one can append an adjectival phrase to a common noun and obtain
a new common noun. Furthermore,  GF is equipped  with a library of mappings from abstract
syntax to the concrete s usyntax of varionatural languages. These mappings can be inverted by GF, thus offering
parsing from natural text into abstract syntax. However, in this project we skip the
parsing phase and use the parse trees constructed by \citet{Ljunglof:2012},
thereby avoiding any syntactic ambiguity. \todo{Do you want to further comment here?}

\paragraph{Coq}
Coq is an interactive theorem prover (proof assistant) based on the
calculus of inductive constructions (CiC), i.e.  a lambda calculus
with dependent types. Coq is arguably one of the leading proof
assistants. It has famously been instrumental in formalising a proof
of the four-color theorem \citep{Gonthier:2008}, a proof of the odd
order theorem \citep{Gonthier:2013}. Coq is a very powerful
reasoning engine that makes it fit for the task of NLI, when the
latter is formalized as theorem proving task. Its underlying higher
order logic is furthermore powerful enough to express fine-grained
logical semantics. Furthermore, Coq supports a number of constructions
that we will be using notably dependent typing and subtyping. These
features allow us to take care of specific NL phenomena
(e.g. subtyping is used in defining different adjectival inference
classes).

\paragraph{Dynamic Monadic Semantics}
Dynamic Monadic Semantics have been proven to be an effective way of
dealing with anaphora and ellipsis. There are a number of approaches
using monads or other equivalent constructions (e.g. continuations as in the work of
\citet{de2006}) for anaphora and ellipsis
\citet{Shan:2002,unger:2011,Barker:04,de2016,charlow:2017}. In this
paper, we follow the approach described by
\citet{bernardy_jolli}. More details on this will be given in the next
section, where the actual system is presented.

@@

Our system consists of two main parts.
\begin{enumerate}
\item A converter from syntax trees to types. The syntax trees follow
  the GF formalism, and the types follow the Coq formalism. The
  converter itself is a Haskell Program, which implements a dynamic
  semantics and comprises the bulk of our system.
\item A number of type-theoretical combinators. These take care of the
  natural semantics which have no influence on the dynamical
  aspects. Such aspects include for example the treatment of adjectives
  (intersective, subsective, etc.) and adverbs (veridical or not).
\end{enumerate}

@@

\paragraph{anaphora} Thanks to the above system, we can handle many
anaphoric cases, including E-Type and Donkey pronouns. Indeed, even
objects which have no syntactic representation can be added to the
environment. We follow here the general pattern outlined by
\citet{unger:2011}. However, we use a more general scope-extension
mechanism, which allows us to support examples such as the following:
\begin{lingex}
\item
    \label{ex:anaphora-scope}
\textbf{P1}	Every committee has a chairman.\\
\textbf{P2}	He is appointed its members.\\
\textbf{H} 	Every committee has a chairman appointed by members of the committee. \\ \textbf{Label:	YES} [FraCaS 122]
\end{lingex}
In the above example, the pronoun ``he'' is allowed to refer to the
object quantified over by ``every'', whose scope is extended
accordingly.

\paragraph{ellipsis} Ellipsis is handled in essentially the same way
as anaphora. This method is made especially straightforward thanks to
using GF syntax tree, which requires an explicit argument for each
predicate. Thus, ellpsis are made explicit by the parsing phase. Such
ellptic expressions are handled in the same way as anaphora.
\begin{lingex}
\item
    \label{ex:acd}
\textbf{P1}	Bill spoke to everyone that John did [elliptic V2].\\
\textbf{P2}	John spoke to Mary.\\
\textbf{Q} 	Did Bill speak to Mary?\\
\textbf{H} 	Bill spoke to Mary. \\ \textbf{Label:	Yes} [FraCas 173]
\end{lingex}
For example, in (FraCaS 173, P1), the argument of ``did'' is
explicitly marked as an elliptic V2, which we resolve to ``speak'' in
that context.
\paragraph{definites} A naive way to handle definites is an
existential (or sigma). However, if the semantics does not feature a
dynamic element, then the existential quantification is introduced
locally. This means that the quantifier can be introduced in the wrong
context. Consider the phrase ``everyone pets the dog''. The structure
of the interpretation would be
$∀x. person(x) → ∃y. dog(y) ∧ pet(x,y)$.

Instead, our take is that definites should be treated as an anaphoric
expression with an implicit referrent. That is, if the referent is not
found in the discourse, then it will be forcibly introduced, using an
existential type, \emph{at the top-level} of the expression. To be
able to do this, we record all definites without referent, using
another portion of the environment (using a monadic effect). For the
above example, we obtain the desired interpretation:
$∃y. dog(y) ∧ (∀x. person(x) → pet(x,y))$.

\paragraph{Phrasal comparatives}
Previous attempts to tackle the section of the FraCaS test suite
devoted to comparatives showed that handling them is not easy. Our
strategy here is to leverage our dynamic semantics, revealing an
anaphoric element of comparatives. Indeed, Consider the hypothesis of
(FraCaS 239): ``ITEL won more orders than APCOM lost.''  We postulate
that this sentence is equivalent to the following two separate parts:
``APCOM lost zero or more orders. ITEL won more orders [than some
elliptic quantity].''  A quantity is introduced every time we talk
about some quantity (indexed by a CN, in this case ``orders''), and it
can be referred to by a comparative, later in the discourse. Using
this idea, we can go one level deeper in the interpretation of our
example: ``APCOM lost θ orders. θ $\ge$ 0.  ITEL won at least θ+1
orders.''. We see here how the quantities are introduced. They are
added to the environment so that, they can be referred to as elliptic
quantity expressions. Finally, ``more'' is intepreted as ``at least
<elliptic quantity>+1''. This treatment, which we illustrated here on
an example, is systematic in our implementation.

\paragraph{Adjectives}
In our system, we interpret gradable adjectives using a pair of a
measure $m : objects → Z$ and and a threshold $τ : Z$, where $Z$ is
treated as an abstract ordered ring by Coq. (This structure has no
dynamic aspect in our model, and thus is entirely handled within Coq.)
For subsective adjectives, $τ$ will additionally depend on the class
of the object in question. This structure has the benefit that
opposite adjectives can be easily represented (measures are opposites
$∀x. m_1 x = -m_2 x$ and thresholds do not overlap $τ_1 + τ_2 >
0$). Formalization aside, this idea is reminiscent of degree-based
approaches to gradable adjectives of
\citet{Cresswell:1976,Kennedy:2007}.  Additionally,
adjectival phrases, as present in the FraCaS suite, are interpreted as
linear inequations in $Z$. Solving systems of such inequations is
decidable. Indeed, the tactic that Coq offers for this purpose can
solve all such problems in the FraCaS suite, automatically.

@@

FraCaS intends to classify each problem as either entailment (YES),
entailment of the opposite (NO) or no entailment (UNK).  In this work,
we have amended the FraCaS suite to correct a few problems. First,
certain test case are not formed correctly. Those were already
identified by Bill MacCartney\todo{cite} as such (using an ``undef''
labelling), and we removed those. Second, a few test cases occur twice
in the suite, but with two different labelling (one YES and one UNK),
with an annotation that those labellings correspond to different
readings. However, elsewhere in the suite, if a problem has several
readings but only one has entailment, it occurs only once and is
marked as YES. To make the test suite consistent, if one reading
yields entailment we have always considered it as YES. We have also
removed case 199 (which appears to be vacuous). Finally we changed the
labelling of 6 cases 
% the cases listed in Table \ref{tab:overrulings},
which appeared to have been
misclassified (as justified in supplementary material). We note that the majority of the mistaken
classifications occur in sections 3 and 4, which have not been
previously attempted and thus, we propose, have not been properly
scrutinized.
% \begin{table}
%   \centering
%   \small
% \begin{tabularx}{\columnwidth}{ccp{3.8cm}}
%  test case & new class & comment \\ \hline
%     005 & UNK & missing hypothesis: there are italian tenors \\
%    056 &  Yes & Already identified as such by MacCartney \\
%   069 & Unk & Mary could have used someone else's workstation \\
%   119 & Unk & \textit{ibid.} \\
%   181 & Yes & for the same reason as 180 \\
%   226 & Yes &
% \end{tabularx}
%   \caption{Overruled FraCaS cases}
%   \label{tab:overrulings}
% \end{table}

@@

\section{Future work}
\section{Conclusion}

>>>>

\begin{abstract}
 In this paper, we present a system for Natural Language Inference which uses a Haskell dynamic semantics  converter from abstract syntax trees to Coq types. The system makes use of the fine-grainedness of a dynamic semantics system with the powerfulness of a state-of-the-art proof assistant, like Coq. We evaluate the system on all sections of the FraCaS test suite, excluding section 6. This is the first system that does a complete run on the anaphora and ellipsis sections of the FraCaS and has the best accuracy overall when compared with previous systems. 
\end{abstract}
\section{Introduction}
Natural Language Inference (NLI) is the task of determining of whether
an NL hypothesis H follows from an NL premise(s) P. NLI has received a
lot of attention in the Computational Semantics literature and has
been approached using a variety of techniques, ranging from logical
approaches
\citep{bos:2008,Mineshima:2015,Abzianidze:2015,bernardy_type_2017},
all the way to the recent Deep Learning (DL) models for NLI. The
latter approaches, following a general trend in NLP, have been
dominating NLI and a number of impressive results have been
produced \citep{kim:2018,radford:2018,liu:2019}.\footnote{These are
  the three systems with the best results on SNLI in increasing
  order at the time of writing.}  State-of-the-art DL systems achieve an accuracy of around
0.9 when tested on suitable datasets. However, the datasets that are
used are assuming a definition of inference that can be thought to be
`looser' or less precise compared to the definition assumed in
platforms based in logical approaches. For example, consider the following
 example  from the SNLI dataset, predominatly used to test DL
approaches:
%
\begin{lingex}
\item % An SNLI example \\
  \label{ex:snli}
\textbf{P}  A  man selling donuts to a customer during a world
exhibition event held in the city of Angeles. \\
 \textbf{H}  A woman drinks her coffee in a small cafe.\\
\textbf{Label: Contradiction} [SNLI]
% \item % An example from FraCaS\\
% \textbf{P1} A Scandinavian won the Nobel Prize.\\
% \textbf{P2}	Every Swede is  Scandinavian.\\
% \textbf{Q}  Did a Swede win the Nobel prize? \\
% \textbf{H} A Swede won the Nobel prize.\\
% \textbf{Label: Yes} [FraCaS 173]
\end{lingex}
%
In \ref{ex:snli}, a number of non-trivial
assumptions have to be made in order to arrive at a contradiction: a) the two
situations described have to be taken to refer to the same situation
in order to judge that the latter  contradicts the former, b)
the indefinite article in the premise has to be identified with the
indefinite article in the hypothesis. (Additionally considering that a
person cannot be a man selling donuts and a woman drinking coffee at
the same time.) While this can be part of the reasoning humans
perform, it is not the only possibility. More precise, logical
reasoning is also a possibility, and will render the above label as unknown. Furthermore, reasoning can get very fine-grained 
as the Contained Deletion ellipsis example \ref{ex:acd} below shows:

\begin{lingex}
\item
    \label{ex:acd}
\textbf{P1}	Bill spoke to everyone that John did [elliptic V2].\\
\textbf{P2}	John spoke to Mary.\\
\textbf{Q} 	Did Bill speak to Mary?\\
\textbf{H} 	Bill spoke to Mary. \\ \textbf{Label:	Yes} [FraCas 173]
\end{lingex}

@@

\paragraph{GF}
In GF, abstract syntax is comprised of: a) a number of syntactic
categories, and b) a number of syntactic construction functions. The latter
provide the means to compose basic syntactic categories into more complex
ones.  For example, consider the constructor: $AdjCN : AP → CN → CN$. This  expresses
that one can append an adjectival phrase to a common noun and obtain
a new common noun. Furthermore,  GF is equipped  with a library of mappings from abstract
syntax to the concrete syntax of various natural languages.
% These mappings can be inverted by GF, thus offering parsing from natural text into abstract syntax.
However, in this project we skip the
parsing phase and use the parse trees constructed by \citet{Ljunglof:2012},
thereby avoiding any syntactic ambiguity.

\paragraph{Coq}
Coq is an interactive theorem prover (proof assistant) based on the
calculus of inductive constructions (CiC), i.e.  a lambda calculus
with dependent types. Coq is  one of the leading proof
assistants. It has famously been instrumental in formalising a proof
of the four-color theorem \citep{Gonthier:2008}, a proof of the odd
order theorem \citep{Gonthier:2013}. Coq is a very powerful
reasoning engine that makes it fit for the task of NLI, when the
latter is formalized as theorem proving task. It involves a very higher order logic  and supports a number of constructions, notably dependent typing and subtyping, that are instrumental in expressing NL semantics.

\paragraph{Dynamic Monadic Semantics}
Dynamic Monadic Semantics have been proven to be an effective way of
dealing with anaphora and ellipsis. There are a number of approaches
using monads or other equivalent constructions (e.g. continuations as in the work of
\citet{de2006}) for anaphora and ellipsis
\citet{Shan:2002,unger:2011,Barker:04,de2016,charlow:2017}. In this
paper, we follow the approach described in
\citet{bernardy_jolli}. More details on this will be given in the next
section.

@@

Our system consists of two main parts.
\begin{enumerate}
\item A converter from syntax trees to types. The syntax trees follow
  the GF formalism, and the types follow the Coq formalism. The
  converter itself is a Haskell Program, which implements a dynamic
  semantics and comprises the bulk of our system.
\item A number of type-theoretical combinators, that encode aspects of NL semantics which have no influence on the dynamic
  part. Such aspects include, e.g.,  the treatment of adjectives
  (intersective, subsective, etc.) and adverbs (veridical or not).
\end{enumerate}

@@

\paragraph{anaphora} Thanks to the above system, we can handle many
anaphoric cases, including E-Type and Donkey pronouns. Indeed, even
objects which have no syntactic representation can be added to the
environment. We follow here the general pattern outlined by
\citet{unger:2011}. However, we use a more general scope-extension
mechanism, which allows us to support examples like  the following:
\begin{lingex}
\item
    \label{ex:anaphora-scope}
\textbf{P1}	Every committee has a chairman.\\
\textbf{P2}	He is appointed its members.\\
\textbf{H} 	Every committee has a chairman appointed by members of the committee. \\ \textbf{Label:	YES} [FraCaS 122]
\end{lingex}
In the above example, the pronoun ``he'' is allowed to refer to the
object quantified over by ``every'', whose scope is extended
accordingly.

\paragraph{ellipsis} Ellipsis is handled in essentially the same way
as anaphora. This method is made especially straightforward thanks to
using GF syntax trees, which require an explicit argument for each
predicate. Thus, ellipsis are made explicit by the parsing phase. Such
ellptic expressions are handled in the same way as anaphora.

For example, in \ref{ex:acd}, the argument of ``did'' is
explicitly marked as an elliptic V2, which we resolve to ``speak'' in
that context.
\paragraph{definites} A naive way to handle definites is an
existential (or sigma). However, if the semantics does not feature a
dynamic element, then the existential quantification is introduced
locally. This means that the quantifier can be introduced in the wrong
context. Consider the phrase ``everyone pets the dog''. The structure
of the interpretation would be
$∀x. person(x) → ∃y. dog(y) ∧ pet(x,y)$.

Instead, our take is that definites should be treated as an anaphoric
expression with an implicit referrent. That is, if the referent is not
found in the discourse, then it will be forcibly introduced, using an
existential type, \emph{at the top-level} of the expression. To be
able to do this, we record all definites without referent, using
another portion of the environment (using a monadic effect).
%SC
%For the
%above example, we obtain the desired interpretation:
%$∃y. dog(y) ∧ (∀x. person(x) → pet(x,y))$.

\paragraph{Phrasal comparatives}
Previous attempts to tackle the section of the FraCaS test suite
devoted to comparatives showed that handling them is not easy. Our
strategy here is to leverage our dynamic semantics, revealing an
anaphoric element of comparatives. Indeed, Consider the hypothesis of
(FraCaS 239): ``ITEL won more orders than APCOM lost.''  We postulate
that this sentence is equivalent to the following two separate parts:
``APCOM lost zero or more orders. ITEL won more orders [than some
elliptic quantity].''  A quantity is introduced every time we talk
about some quantity (indexed by a CN, in this case ``orders''), and it
can be referred to by a comparative, later in the discourse. Using
this idea, we can go one level deeper in the interpretation of our
example: ``APCOM lost θ orders. θ $\ge$ 0.  ITEL won at least θ+1
orders.''. We see here how the quantities are introduced. They are
added to the environment so that, they can be referred to as elliptic
quantity expressions.
%SC
%Finally, ``more'' is intepreted as ``at least
%<elliptic quantity>+1''. This treatment, which we illustrated here on
%an example, is systematic in our implementation.

\paragraph{Adjectives}
In our system, we interpret gradable adjectives using a pair of a
measure $m : objects → Z$ and and a threshold $τ : Z$, where $Z$ is
treated as an abstract ordered ring by Coq. (This structure has no
dynamic aspect in our model, and thus is entirely handled within Coq.)
For subsective adjectives, $τ$ will additionally depend on the class
of the object in question. This structure has the benefit that
opposite adjectives can be easily represented (measures are opposites
$∀x. m_1 x = -m_2 x$ and thresholds do not overlap $τ_1 + τ_2 >
0$). Formalization aside, this idea is reminiscent of degree-based
approaches to gradable adjectives of
\citet{Cresswell:1976,Kennedy:2007}.  Additionally,
adjectival phrases, as present in the FraCaS suite, are interpreted as
linear inequations in $Z$. Solving systems of such inequations is
decidable.
% SC
%Indeed, the tactic that Coq offers for this purpose can
%solve all such problems in the FraCaS suite, automatically.

@@

FraCaS intends to classify each problem as either entailment (YES),
entailment of the opposite (NO) or no entailment (UNK).  In this work,
we have amended the FraCaS suite to correct a few problems. First,
certain test case are not formed correctly. Those were already
identified by Bill MacCartney \citet{maccartney:2007} as such (using an ``undef''
labelling), and we removed those. Second, a few test cases occur twice
in the suite, but with two different labelling (one YES and one UNK),
with an annotation that those labellings correspond to different
readings. However, elsewhere in the suite, if a problem has several
readings but only one has entailment, it occurs only once and is
marked as YES. To make the test suite consistent, if one reading
yields entailment we have always considered it as YES. We have also
removed case 199 (which appears to be vacuous). Finally we changed the
labelling of 6 cases
%(SC: Shall we remove this to the supplementary material and just say to have a look at how we amended the suite?)
% the cases listed in Table \ref{tab:overrulings},
which appeared to have been
misclassified (as justified in supplementary material). We note that the majority of the mistaken
classifications occur in sections 3 and 4, which have not been
previously attempted and thus, we propose, have not been properly
scrutinized.
% \begin{table}
%   \centering
%   \small
% \begin{tabularx}{\columnwidth}{ccp{3.8cm}}
%  test case & new class & comment \\ \hline
%     005 & UNK & missing hypothesis: there are italian tenors \\
%    056 &  Yes & Already identified as such by MacCartney \\
%   069 & Unk & Mary could have used someone else's workstation \\
%   119 & Unk & \textit{ibid.} \\
%   181 & Yes & for the same reason as 180 \\
%   226 & Yes &
% \end{tabularx}
%   \caption{Overruled FraCaS cases}
%   \label{tab:overrulings}
% \end{table}

@@

\section{Conclusions and Future Work}
